---
title: "Weather code"
author: "Lara Zini & Adrian Correndo"
date: "2022-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 00. Packages weather
```{r warning= F, message=F}
#install.packages("easypackages")
library(easypackages)
libraries('tidyverse') # Data wrangling
libraries('lubridate') # Dates operations
libraries('kableExtra') # Table formatting
libraries('openxlsx') # Open excel files
libraries('daymetr','chirps','nasapower') # Weather databases
libraries('vegan') # Shannon Diversity Index
libraries("ggplot2","ggrepel","patchwork")
packages("tidybayes")
```

## 01. DAYMET function

Here we download the daily-weather data from the DAYMET database, and we process it to obtain common variables of agronomic value.

```{r}
# Function
weather.daymet <- function(input, dpp=0){ input %>%
  mutate(Weather = pmap(list(ID = ID,
                             lat = latitude,
                             lon = longitude,
                             sta = Start - dpp,
                             end = End),
                        
# Retrieving daymet data
        function(ID,lat,lon,sta,end){
                download_daymet(site = ID,
                                lat = lat, lon = lon,
                                # Extracting year
                                start = as.numeric(substr(sta,1,4)),
                                end = as.numeric(substr(end,1,4)),
                                internal = T, simplify = T)})) %>% 
    
# Organizing dataframe (Re-arranging rows and columns)
  mutate(Weather = Weather %>% 
# Adjusting dates format
           map(~mutate(.,
                       Date = as.Date(as.numeric(yday)-1, # Day of the year
                                         origin = paste0(year,'-01-01')),
                       Year = year(Date),
                       Month = month(Date),
                       Day = mday(Date)))  %>% 
           
           map(~dplyr::select(., yday, Year, Month, Day, Date,
                              measurement,value)) %>% 
           map(~spread(., 'measurement','value'))  %>% 
# Renaming variables
           map(~rename_all(., ~c("DOY", # Date as Day of the year
                                 "Year", # Year
                                 "Month", # Month 
                                 "Day", # Day of the month
                                 "Date", # Date as normal format
                                 "DL", # Day length (sec)
                                 "PP", # Precipitation (mm)
                                 "Rad", # Radiation (W/m2)
                                 "SWE", # Snow water (kg/m2)
                                 "Tmax", # Max. temp. (degC)
                                 "Tmin", # Min. temp. (degC)
                                 "VPD"))))  %>% # Vap Pres Def (Pa)
# Processing data given start and ending dates
  mutate(Weather = pmap(list(sta=Start-dpp,
                             end = End,data=Weather), # Requested period
                        #~filter(..3, Date>=..1 & Date<= ..2))) %>% unnest() %>%
                        function(sta, end, data){
                          filter(data, Date >= sta & Date <= end) 
                        } )) %>% unnest(cols = c(Weather)) %>% 
  
# Converting units or adding variables
  mutate(Rad = Rad*0.000001*DL, # Radiation (W/m2 to MJ/m2)
         Tmean = (Tmax+Tmin)/2, # Mean temperature (degC),
         VPD = VPD / 1000, # VPD (Pa to kPa),
         # Creating variables for ET0 estimation
         lat_rad = latitude*0.0174533,
         dr = 1 + 0.033*cos((2*pi/365)*DOY),
         Sd = 0.409*sin((2*pi/365)*DOY - 1.39),
         ws = acos(-tan(lat_rad)*tan(Sd)),
         Ra = (24*60)/(pi) * Gsc * dr * (ws*sin(lat_rad)*sin(Sd)+
                                  cos(lat_rad)*sin(ws)),
         ET0_HS = 0.0135 * kRs * (Ra / 2.45) * (sqrt(Tmax-Tmin)) * (Tmean + 17.8),
         DL = (DL/60)/60 # Day length (hours)
          ) %>% dplyr::select(-lat_rad,-dr,-Sd,-ws,-Ra)
}


### INTERVALS

# Defining function to summarize historical weather (years)
historical.years <- function(hist.data) {
  
# By year
hist.data %>% group_by(ID,Year) %>% 
  dplyr::select(ID, latitude, longitude, Start, End, Date, Year, Month) %>% 
  summarise(Start.in = min(Date),
            End.in = max(Date))

}

# Defining function to summarize historical weather (years & months)
historical.years.months <- function(hist.data) {
  
# By year*month
hist.data %>% group_by(ID, Year, Month) %>% 
  dplyr::select(ID, latitude, longitude, Start, End, Date, Year, Month) %>% 
  summarise(Start.in = min(Date),
            End.in = max(Date))

}

```

## 02. Summary function - DAYMET & NASA-POWER
```{r}

# Defining the function to summarize DAYMET and/or NASA-POWER
summary.daymet.nasapower <- function(input, intervals) {

  intervals %>% 
  
# Merging weather data
  left_join(input %>% 
# Nesting weather data back for each site-ID
  dplyr::select_if(
    names(.) %in% c("ID", "Crop", "Site", "Date","DL", "PP",
                    "Rad", "Tmax", "Tmin", "Tmean", "VPD", "ET0_HS")) %>% 
  group_by(ID) %>% nest(.key = 'Weather') %>% ungroup()) %>% 
  
  mutate(Weather = pmap(list(x = Start.in,y = End.in, data = Weather),
                        function(x, y, data){
                          filter(data, Date >= x & Date < y)} ) )  %>% 
  
  mutate(Weather = Weather %>% # User must addapt depending on the crop
           map(~mutate(.,
                # Ext. Prec. event
                EPEi = case_when(PP>25~1, TRUE~0),
                # Ext.Temp. event
                ETEi = case_when(Tmax >= 33~1, TRUE~0), 
                # Tmax factor,  crop heat units (CHU)
                Ymax = case_when(Tmax < 10~0, 
                            TRUE ~ 3.33*(Tmax-10)-0.084*(Tmax-10)),
                # Tmin factor, Crop heat units (CHU)
                Ymin = case_when(Tmin<4.44~0, 
                            TRUE ~ 1.8*(Tmin-4.44)), 
                # Daily CHU
                Yavg = (Ymax+Ymin)/2,
                # Tmin threshold Growing Degrees.
                Gmin = case_when(Tmin >= 10 ~ Tmin, 
                            TRUE ~ 10),
                # Tmax threshold Growing Degrees.
                Gmax = case_when(Tmax <= 33 ~ Tmax, 
                            TRUE ~ 33),
                # Daily Growing Degree Units.
                GDU = ((Gmin + Gmax)/2) - 10
                       ) ) ) %>% 

# Summary for each variable

  mutate(# Duration of interval (days)
         Dur = Weather %>% map(~nrow(.)),
         # Accumulated PP (mm)
         PP = Weather %>% map(~sum(.$PP)),
         # Mean Temp (C)
         Tmean = Weather %>% map(~mean(.$Tmean)),
         # Accumulated Rad (MJ/m2)
         Rad = Weather %>% map(~sum(.$Rad)),
         # Accumulated VPD (kPa)
         VPD = Weather %>% map(~sum(.$VPD)),
         # Accumulated ET0 (mm)
         ET0_HS = Weather %>% map(~sum(.$ET0_HS)),
         # Number of ETE (#)
         ETE = Weather %>% map(~sum(.$ETEi)),
         # Number of EPE (#)
         EPE = Weather %>% map(~sum(.$EPEi)),
         # Accumulated Crop Heat Units (CHU)
         CHU = Weather %>% map(~sum(.$Yavg)),
         # Shannon Diversity Index for PP
         SDI = Weather %>% map(~diversity(.$PP, index="shannon")/
                                 log(length(.$PP))),
         # Accumulated Growing Degree Days (GDD) 
         GDD =  Weather %>% map(~sum(.$GDU))) %>% 
  
# Additional indices and final units
  dplyr::select(-Weather) %>% unnest() %>% 
  mutate(# Photo-thermal quotient (Q)
         Q_chu = Rad/CHU,
         Q_gdd = Rad/GDD,
         # Abundant and Well Distributed Water
         AWDR = PP*SDI) 
}
```

## 03 Weather DigitalAg _ VER MESES EN FILTER!!
```{r warning=F, message=F}
# Open file from excel if you have
# data_table <- openxlsx::read.xlsx(xlsxFile = "my_file.xlsx", sheet = 1)

# Create data frame with DigitalAg coordinates
my_data <- 
  tibble::new_tibble(
  list(ID = "DigitalAg",
       latitude = 39.216948, 
       longitude = -96.596674,
       Start = rep('1982-01-01',),
       End = rep('2021-12-31',))) %>% 
  mutate_at(vars(c("Start","End")), ~as.Date(., format='%Y-%m-%d'))

  

# Constants for ET0 (Cobaner et al., 2017)
# Solar constant
Gsc = 0.0820 # (MJ m-2 min-1)
# Radiation adjustment coefficient (Samani, 2004)
kRs = 0.17

# Path
path = paste0(getwd(), '/') # Current directory or any path.

# Specify input = dataframe containing historical dates from sites 

weather.daymet(input = my_data) -> hist.weather.DigitalAg


# Exporting data as a .csv file
# write.csv(hist.weather.DigitalAg, row.names = F, na='', file = paste0(path, 'hist.weather.DigitalAg.csv'))


# Specify hist.data = dataframe containing the historical weather data to summarize
years = historical.years(hist.data = hist.weather.DigitalAg)

# Specify hist.data = dataframe containing the historical weather data to summarize
years.months = historical.years.months(hist.data = hist.weather.DigitalAg)

# input = dataframe containing the historical weather data.
# intervals = type of historical intervals (years, years.months)

# Summarizing historical weather
historical.summary.DigitalAg <-
  summary.daymet.nasapower(input = hist.weather.DigitalAg,
                           intervals = years.months) %>%
  # Filter 
dplyr::filter(Month %in% c(1:12)) 
  

# Creating a table to visualize data
kbl(historical.summary.DigitalAg) %>%
  kable_styling(font_size = 7, position = "center", latex_options = c("scale_down"))

# Exporting data as a .csv file
# Daymet
# write.csv(historical.summary.DigitalAg, row.names = F, na='',
#          file = paste0(path, 'Historical_summary_DigitalAg.csv'))

#View(historical.summary.DigitalAg)

# PP & Temp summaries
DigitalAg.hist.PP.Temp <- historical.summary.DigitalAg %>% 
  group_by(ID, Year) %>% 
  summarise(PP = sum(PP), 
            Tmean = mean(Tmean),
            ETE = sum(ETE),
            EPE = sum(EPE)) %>%
    mutate(PP_met = PP,
           Tmean_met = Tmean,
           PP = PP / 25.4,
           Tmean = Tmean * 1.8 + 32)
DigitalAg.hist.PP.Temp

```


##COMPLETEYEAR_DATA
```{r}
DigitalAg.hist.PP.Temp <-
  DigitalAg.hist.PP.Temp %>% 
  group_by(ID) %>% 
  mutate(PP_mean = mean(PP),
         T_mean = mean(Tmean),
         PP_anom = (PP*100/PP_mean)-100, 
         T_anom = Tmean-T_mean,
         labyr = case_when(PP_anom < -10 & T_anom < -.5 ~ "Dry & Cool",
                           PP_anom < -10 & T_anom > .5 ~ "Dry & Warm",
                           PP_anom > 10 & T_anom < -.5 ~ "Wet & Cool",
                           PP_anom > 10 & T_anom > .5 ~ "Wet & Warm",
                           between(PP_anom,-10,10) & T_anom < -.5 ~ "Normal & Cool",
                           between(PP_anom,-10,10) & T_anom > .5 ~ "Normal & Warm",
                           PP_anom < -10 & between(T_anom,-.5,.5) ~ "Dry & Normal",
                           PP_anom > 10 & between(T_anom,-.5,.5) ~ "Wet & Normal",
                           between(PP_anom,-10,10) & between(T_anom,-.5,.5) ~ "Normal"),
                 symb_point = ifelse(labyr != "Normal", "yes", "no") ) %>% 

ungroup()
    
DigitalAg.hist.PP.Temp
```



#COMPLETE YEAR_ ANOMALY PLOT _ WEATHER ANOMALIES
```{r}

plot_COMPLETE_anom_10 <- ggplot(data= DigitalAg.hist.PP.Temp)+ 
    geom_point(aes(x=T_anom, y = PP_anom ), shape = 16, color= "black", size = 1)+
    geom_hline(yintercept = 0, linetype = "dashed")+
    geom_vline(xintercept = 0, linetype = "dashed")+
    labs(x='Temp_Anom', y= 'Precip_Anom')+ 
    ggtitle("Weather anomalies - 1982 to 2021")+ #PRESTAR ATENCION TITULOS
    geom_label(aes(x = -2.2, y = 60, 
          label = "Wet & Cool"), 
          stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE SUP-IZQ
    geom_label(aes(x = -2.2, y = -60,
         label = "Dry & Cool"),
          stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE INF-IZQ
    geom_label(aes(x = 2.5, y = -60,
           label = "Dry & Warm"),
           stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE INF-DER
    geom_label(aes(x = 2.5, y = 60, 
               label = "Wet & Warm"),
           stat = "unique") + #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE SUP-DER
    geom_text_repel(aes(x=T_anom, y = PP_anom ,
                    label = as.integer(Year),label.size = 0.25)) + 
    labs(color='Year', caption = "T_mean= 54.8 ºF | PP_mean= 34.02 in")+
    scale_x_continuous(limits=c(-3,3), breaks = seq(-3,3, by = 1))+
     scale_y_continuous(limits=c(-60,60), breaks=seq(-60,60, by = 20))+
theme_bw(base_size = 12, base_family = "Verdana", base_line_size = 0.5, base_rect_size = 0.5)+
    theme(panel.grid.major = element_blank())
    
plot_COMPLETE_anom_10
    

##GUARDAR

ggsave(plot = plot_COMPLETE_anom_10,
       filename = paste0('../DigitalAg_Field_Project/output/V2/',
                         'plot_COMPLETE_anom_10','.png'),
       device='png',
       width = 23,height = 17,units = 'cm')



```

##COMPLETE_ EXTREME PP PLOT
```{r}

# NO USO!!
### PLOT DE ExTE _ CANTIDAD DE DIAS QUE SUPERAN CIERTA TEMPERATURA (arriba se calculo ETE y EPE, se determino +30 grados centigrados y lluvias de + de 20 mm)

#plot_COMPLETE_EXTTEMP <- DigitalAg.hist.PP.Temp %>% 
  #group_by(ID) %>% 
 # mutate(ETE_av = mean(ETE)) %>% ungroup() %>% 
 # ggplot(aes(x=Year, y = ETE))+
 # geom_col()+
 # geom_hline(aes(yintercept = ETE_av))+
 #   labs(x='Year', y= 'Number of extreme temperature')+
 #   ggtitle ("Extreme temperature - 1982 to 2021")
#
#ggsave(plot = plot_COMPLETE_EXTTEMP,
     #  filename = paste0('../DigitalAg_Field_Project/output/V2/',
                        # ' plot_COMPLETE_EXTTEMP','.png'),
      # device='png',
     #  width = 17,height = 10,units = 'cm')



## PLOT ExPP     

plot_COMPLETE_EXTPP <-DigitalAg.hist.PP.Temp %>% 
  group_by(ID) %>% 
  mutate(EPE_av = mean(EPE)) %>% ungroup() %>% 
  ggplot(aes(x=Year, y = EPE))+
  geom_col()+
  geom_hline(aes(yintercept = EPE_av))+
    labs(x='Year', y= 'Number of extreme precipitation')+
    ggtitle ("Extreme precipitation - 1982 to 2021", subtitle = "ExPP: Number of days that exceed 1 in")+
    theme_bw(base_size = 12, base_family = "Verdana", base_line_size = 0.5, base_rect_size = 0.5)+
    theme(panel.grid.major = element_blank())

plot_COMPLETE_EXTPP

#GUARDAR 
ggsave(plot = plot_COMPLETE_EXTPP,
       filename = paste0('../DigitalAg_Field_Project/output/V2/',
                         ' plot_COMPLETE_EXTPP','.png'),
       device='png',
       width = 25,height = 15,units = 'cm')

```










##SUMMER_DATA ANOMALY 
```{r}
DigitalAg.hist.PP.Temp <-
  DigitalAg.hist.PP.Temp %>% 
  group_by(ID) %>% 
  mutate(PP_mean = mean(PP),
         T_mean = mean(Tmean),
         PP_anom = (PP*100/PP_mean)-100, 
         T_anom = Tmean-T_mean,
         labyr = case_when(PP_anom < -10 & T_anom < -.5 ~ "Dry & Cool",
                           PP_anom < -10 & T_anom > .5 ~ "Dry & Warm",
                           PP_anom > 10 & T_anom < -.5 ~ "Wet & Cool",
                           PP_anom > 10 & T_anom > .5 ~ "Wet & Warm",
                           between(PP_anom,-10,10) & T_anom < -.5 ~ "Normal & Cool",
                           between(PP_anom,-10,10) & T_anom > .5 ~ "Normal & Warm",
                           PP_anom < -10 & between(T_anom,-.5,.5) ~ "Dry & Normal",
                           PP_anom > 10 & between(T_anom,-.5,.5) ~ "Wet & Normal",
                           between(PP_anom,-10,10) & between(T_anom,-.5,.5) ~ "Normal"),
                 symb_point = ifelse(labyr != "Normal", "yes", "no") ) %>% 

ungroup()%>%
    

##PARA SUMMER CROPS / creo estas columnas para resaltar años extremos en SUMMER
mutate(T_anom_ext = case_when(T_anom < -2.5 | T_anom > 2.5 ~ "yes", TRUE ~ "no"), 
PP_anom_ext = case_when(PP_anom < -30 | PP_anom > 30 ~ "yes", TRUE ~ "no"), 
year_ext = case_when(PP_anom_ext == "yes" | T_anom_ext == "yes" ~ "yes",TRUE ~ "no"))


#View(DigitalAg.hist.PP.Temp)



```







#SUMM_ ANOMALYPLOT _ 10 AÑOS
```{r}

# Plot CON LOS PUNTOS DE LOS ULTIMOS 10 AÑOS RESALTADOS


plot_summer_anom_10 <- ggplot(data= DigitalAg.hist.PP.Temp)+ 
geom_point(aes(x=T_anom, y = PP_anom), shape = 16, size = 1.25)+
    geom_hline(yintercept = 0, linetype = "dashed")+
    geom_vline(xintercept = 0, linetype = "dashed")+
    labs(x='Temp_Anom', y= 'Precip_Anom')+ 
ggtitle("Weather anomalies", #PRESTAR ATENCION TITULOS
       subtitle = "1982 to 2021 (Apr to Sept)")+ #PRESTAR ATENCION TITULOS
geom_label(aes(x = -2.5, y = 50, 
          label = "Wet & Cool"), 
          stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE SUP-IZQ
geom_label(aes(x = -2.5, y = -65,
         label = "Dry & Cool"),
          stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE INF-IZQ
geom_label(aes(x = 2.5, y = -65,
           label = "Dry & Warm"),
           stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE INF-DER
geom_label(aes(x = 2.5, y = 50, 
               label = "Wet & Warm"),
           stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE SUP-DER
geom_point(data = DigitalAg.hist.PP.Temp %>%
    dplyr::filter(Year %in% c(2012:2021)), #FILTRO Y RESALTO LOS ULTIMOS 10 AÑOS
    aes(x=T_anom, y = PP_anom, fill = as.factor(Year)),
          shape = 21, size = 3) + 
    labs(color='Year')+
     labs(fill = "Last 10 years",  caption = "T_mean= 69.8 ºF | PP_mean= 24.8 in")+
    scale_x_continuous(limits=c(-4,4), breaks = seq(-4,4, by = 1))+
     scale_y_continuous(limits=c(-70,70), breaks=seq(-70,70, by = 35))+

    theme_bw(base_size = 12, base_family = "Verdana", base_line_size = 0.5, base_rect_size = 0.5)+
    theme(panel.grid.major = element_blank())
   
plot_summer_anom_10 
    

##GUARDAR

ggsave(plot = plot_summer_anom_10,
       filename = paste0('../DigitalAg_Field_Project/output/V2/',
                         'plot_summer_anom_10','.png'),
       device='png',
       width = 25,height = 15,units = 'cm')



```





#SUM _ type of year _ años extremos.      
```{r}

plot_summer_anom_extrem <- ggplot(data= DigitalAg.hist.PP.Temp)+
geom_point(aes(x=T_anom, y = PP_anom), shape = 16, size = 1.25)+
geom_hline(yintercept = 0, linetype = "dashed")+
geom_vline(xintercept = 0, linetype = "dashed")+
    labs(x='Temp_Anom', y= 'Precip_Anom')+ 
ggtitle("Weather anomalies", #PRESTAR ATENCION TITULOS
          subtitle = "1982 to 2021 (Apr to Sept)")+ #PRESTAR ATENCION TITULOS
geom_label(aes(x = -2.5, y = 50, 
          label = "Wet & Cool"), 
          stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE SUP-IZQ
geom_label(aes(x = -2.5, y = -65,
         label = "Dry & Cool"),
          stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE INF-IZQ
geom_label(aes(x = 2.5, y = -65,
           label = "Dry & Warm"),
           stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE INF-DER
geom_label(aes(x = 2.5, y = 50, 
               label = "Wet & Warm"),
           stat = "unique")+ #P/CAMBIAR DE POSICION ETIQ. DE CUADRANTE SUP-DER
geom_point(data = DigitalAg.hist.PP.Temp %>%
    dplyr::filter(year_ext %in% c("yes")), #FILTRO Y RESALTO LOS AÑOS EXTREMOS
      aes(x=T_anom, y = PP_anom, fill = as.factor(Year)),
          shape = 21, size = 3)+
    labs(fill = "Extreme years",  caption = "T_mean= 69.8 ºF | PP_mean= 24.8 in")+
        scale_x_continuous(limits=c(-4,4), breaks = seq(-4,4, by = 1))+
     scale_y_continuous(limits=c(-70,70), breaks=seq(-70,70, by = 35))+

    theme_bw(base_size = 12, base_family = "Verdana", base_line_size = 0.5, base_rect_size = 0.5)+
    theme(panel.grid.major = element_blank())


plot_summer_anom_extrem  

##guardar

ggsave(plot = plot_summer_anom_extrem,
       filename = paste0('../DigitalAg_Field_Project/output/V2/',
                         'plot_summer_anom_extrem','.png'),
       device='png',
       width = 25,height = 15,units = 'cm')

```


#Sum_ EXTREME PLOT - TEMPERATURE: ONLY July and August - PRECIPITATION APRIL TO SEPT
```{r}

# PLOT DE ExTE _ CANTIDAD DE DIAS QUE SUPERAN CIERTA TEMPERATURA (arriba se calculo ETE y EPE, se determino +33 grados centigrados y lluvias de + de 20 mm)

plot_JulyAug_EXTTEMP <- DigitalAg.hist.PP.Temp %>% 
  group_by(ID) %>% 
  mutate(ETE_av = mean(ETE)) %>% ungroup() %>% 
  ggplot(aes(x=Year, y = ETE))+
  geom_col()+
  geom_hline(aes(yintercept = ETE_av))+
    labs(x='Year', y= 'ExTE')+
    ggtitle ("Extreme temperature - 1982 to 2021 (Jul-Aug)", subtitle = "ExTE: Number of days with temperatures greater than 91.4 F")+
theme_bw(base_size = 12, base_family = "Verdana", base_line_size = 0.5, base_rect_size = 0.5)+
    theme(panel.grid.major = element_blank())

plot_JulyAug_EXTTEMP

ggsave(plot = plot_JulyAug_EXTTEMP,
       filename = paste0('../DigitalAg_Field_Project/output/V2/',
                         ' plot_JulyAug_EXTTEMP','.png'),
       device='png',
       width = 25,height = 15,units = 'cm')



## PLOT ExPP  summer april to sept  

plot_summer_EXTPP <-DigitalAg.hist.PP.Temp %>% 
  group_by(ID) %>% 
  mutate(EPE_av = mean(EPE)) %>% ungroup() %>% 
  ggplot(aes(x=Year, y = EPE))+
  geom_col()+
  geom_hline(aes(yintercept = EPE_av))+
    labs(x='Year', y= 'Number of extreme precipitation')+
    ggtitle ("Extreme precipitation - 1982 to 2021 (Apr to Sept)", subtitle = "ExPP: Number of days that exceed 1 in")+
    theme_bw(base_size = 12, base_family = "Verdana", base_line_size = 0.5, base_rect_size = 0.5)+
    theme(panel.grid.major = element_blank())


plot_summer_EXTPP

## guardar

ggsave(plot = plot_summer_EXTPP,
       filename = paste0('../DigitalAg_Field_Project/output/V2/',
                         ' plot_summer_EXTPP','.png'),
       device='png',
       width = 25,height = 15,units = 'cm')

```


## winter_ DATA ANOMALY 
```{r warning=F, message=F}

DigitalAg.hist.PP.Temp <-
  DigitalAg.hist.PP.Temp %>% 
  group_by(ID) %>% 
  mutate(PP_mean = mean(PP),
         T_mean = mean(Tmean),
         PP_anom = (PP*100/PP_mean)-100, 
         T_anom = Tmean-T_mean,
         labyr = case_when(PP_anom < -10 & T_anom < -.5 ~ "Dry & Cool",
                           PP_anom < -10 & T_anom > .5 ~ "Dry & Warm",
                           PP_anom > 10 & T_anom < -.5 ~ "Wet & Cool",
                           PP_anom > 10 & T_anom > .5 ~ "Wet & Warm",
                           between(PP_anom,-10,10) & T_anom < -.5 ~ "Normal & Cool",
                           between(PP_anom,-10,10) & T_anom > .5 ~ "Normal & Warm",
                           PP_anom < -10 & between(T_anom,-.5,.5) ~ "Dry & Normal",
                           PP_anom > 10 & between(T_anom,-.5,.5) ~ "Wet & Normal",
                           between(PP_anom,-10,10) & between(T_anom,-.5,.5) ~ "Normal"),
                 symb_point = ifelse(labyr != "Normal", "yes", "no") ) %>% 

ungroup()%>%
    
##PARA WINTER CROPS / RESALTO LOS ANOS EXTREMOS.
mutate(T_anom_ext = case_when(T_anom < -3 | T_anom > 3 ~ "yes", TRUE ~ "no"), 
   PP_anom_ext = case_when(PP_anom < -50 | PP_anom > 50 ~ "yes", TRUE ~ "no"), 
   year_ext = case_when(PP_anom_ext == "yes" | T_anom_ext == "yes" ~ "yes",TRUE ~ "no"))



#View(DigitalAg.hist.PP.Temp)




```



# WIN _ ANOMALY PLOT 10 AÑOS
```{r}

plot_winter_anom_10 <- ggplot(data= DigitalAg.hist.PP.Temp)+ 
geom_point(aes(x=T_anom, y = PP_anom), shape = 16, size = 1.25)+
geom_hline(yintercept = 0, linetype = "dashed")+
geom_vline(xintercept = 0, linetype = "dashed")+
    labs(x='Temp_Anom', y= 'Precip_Anom')+
ggtitle("Weather anomalies",
          subtitle = "1982 to 2021 (Oct to Apr)")+
geom_label(aes(x = -2.5, y = 40,
                label = "Wet & Cool"),
            stat = "unique")+
geom_label(aes(x = -2.5, y = -50,
                label = "Dry & Cool"),
            stat = "unique")+
geom_label(aes(x = 2.5, y = -50,
                label = "Dry & Warm"),
            stat = "unique")+
geom_label(aes(x = 2.5, y = 40,
                label = "Wet & Warm"),
            stat = "unique")+
   geom_point(data = DigitalAg.hist.PP.Temp %>%
dplyr::filter(Year %in% c(2012:2021)), 
     aes(x=T_anom, y = PP_anom, fill = as.factor(Year)),
shape = 21, size = 3)  + 
labs(fill = "Last 10 years", caption = "T_mean= 41.8 ºF | PP_mean= 12.28 in" )+
   
    scale_x_continuous(limits=c(-4,4), breaks = seq(-4,4, by = 1))+
     scale_y_continuous(limits=c(-70,70), breaks=seq(-70,70, by = 35))+
    theme_bw(base_size = 12, base_family = "Verdana", base_line_size = 0.5, base_rect_size = 0.5)+
    theme(panel.grid.major = element_blank())

 

plot_winter_anom_10

#para agregar texto en el medio del grafico ("text", posicion y el texto)
#     annotate("text", x = 4, y = 70, label = "T_mean= 41.8 ºF | PP_mean= 12.28 in")


##GUARDAR

ggsave(plot = plot_winter_anom_10,
       filename = paste0('../DigitalAg_Field_Project/output/V2/',
                         'plot_winter_anom_10','.png'),
       device='png',
       width = 25,height = 15,units = 'cm') 
    
```


#WIN_ ANOMALYPLOT_ AÑOS EXTREMOS

```{r}

plot_winter_anom_extrem <- ggplot(data= DigitalAg.hist.PP.Temp)+ 
geom_point(aes(x=T_anom, y = PP_anom), shape = 16, size = 1.25)+
geom_hline(yintercept = 0, linetype = "dashed")+
geom_vline(xintercept = 0, linetype = "dashed")+
    labs(x='Temp_Anom', y= 'Precip_Anom')+
ggtitle("Weather anomalies",
          subtitle = "1982 to 2021 (Oct to Apr)")+
geom_label(aes(x = -2.5, y = 40,
                label = "Wet & Cool"),
            stat = "unique")+
geom_label(aes(x = -2.5, y = -40,
                label = "Dry & Cool"),
            stat = "unique")+
geom_label(aes(x = 2.5, y = -40,
                label = "Dry & Warm"),
            stat = "unique")+
geom_label(aes(x = 2.5, y = 40,
                label = "Wet & Warm"),
            stat = "unique")+
  geom_point(data = DigitalAg.hist.PP.Temp %>%
    dplyr::filter(year_ext %in% c("yes")), #FILTRO Y RESALTO LOS AÑOS EXTREMOS
      aes(x=T_anom, y = PP_anom, fill = as.factor(Year)),
          shape = 21, size = 3)+
    labs(fill = "Extreme years", caption = "T_mean= 41.8 ºF | PP_mean= 12.28 in")+
    scale_x_continuous(limits=c(-4,4), breaks = seq(-4,4, by = 1))+ #LIMITE DE LA ESCALA X, Y DE LOS LIMITES CADA CUANTO SE CORTA. EJ CADA 1.
     scale_y_continuous(limits=c(-70,70), breaks=seq(-70,70, by = 35))+ #LIMITES Y DE LOS LIMITES CADA CUANTO CORTA, EJ 35.
    theme_bw(base_size = 12, base_family = "Verdana", base_line_size = 0.5, base_rect_size = 0.5)+
    theme(panel.grid.major = element_blank()) #TAMAÑO DE LETRAS, TIPO DE LETRA, Y GROSOR DE LAS LINEAS X E Y
 
plot_winter_anom_extrem                   
                  
##GUARDAR

ggsave(plot = plot_winter_anom_extrem,
       filename = paste0('../DigitalAg_Field_Project/output/V2/',
                         'plot_winter_anom_extrem','.png'),
       device='png',
       width = 25,height = 15,units = 'cm') 
    
```

#WIN_ extreme precipitarion    
```{r}


# plot expp

plot_winter_EXTPP <-DigitalAg.hist.PP.Temp %>% 
  group_by(ID) %>% 
  mutate(EPE_av = mean(EPE)) %>% ungroup() %>% 
  ggplot(aes(x=Year, y = EPE))+
  geom_col()+
  geom_hline(aes(yintercept = EPE_av))+
labs(x='Year', y= 'Number of extreme precipitation')+
    ggtitle ("Extreme precipitation - 1982 to 2021 (Oct to Apr)", subtitle = "ExPP: Number of days that exceed 1 in")+
    theme_bw(base_size = 12, base_family = "Verdana", base_line_size = 0.5, base_rect_size = 0.5)+
    theme(panel.grid.major = element_blank())


plot_winter_EXTPP

##guardar
ggsave(plot = plot_winter_EXTPP,
       filename = paste0('../DigitalAg_Field_Project/output/V2/',
                         ' plot_winter_EXTPP','.png'),
       device='png',
       width = 25,height = 15,units = 'cm')



```

